---
title: "Project"
output: 
#runtime: shiny
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source_code: https://github.com/rlatimer/Project.git
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(tidyverse)
library(rio)
library(janitor)
library(ggplot2)
library(here)
library(maps)
library(dplyr)
library(gghighlight)
library(plotly)
library(colorspace)
```
Data Source {.sidebar}
==================================
The carbon footprint of food industry varies by country. These data show the CO2 produced  per capita for each food type for 130 countries. These calculations are based on food balance sheets and population data available from the Food and Agriculture Organization of the United Nations (FAO).

These visualizations intend to compare the dietary sources of CO2 of non-animal based products to animal-based products and the differences in the food production across countries.

Animal products
=========================
Column {data-width=650}
-----------------------------------------------------------------------
```{r data}
initial <- import(here("data", "food_carbon_footprint_data.xlsx")) %>% 
  clean_names() %>%
    as_tibble()

subset <- initial %>% 
  mutate(ranking = as.numeric(ranking)) %>% 
  filter(ranking < 7 |country == "average"| country =="Canada"|country =="Japan"
         |country == "Germany"|country =="Mexico"|country =="South Korea"|country =="China")

#line plot work: CO2/person/year produced by country

initial_longer <- subset %>%
  select("ranking", "country","total_animal_products", "total_nonanimal_products", 
         "animal_nonanimal_difference") %>% 
  pivot_longer(cols = 3:5,
               names_to = "product",
               values_to = "CO2_person_year")

nadiff <- subset %>%
  arrange(desc(animal_nonanimal_difference)) %>% 
  select("ranking", "country", "animal_nonanimal_difference") %>% 
  pivot_longer(cols = 3,
               names_to = "product",
               values_to = "CO2_person_year")

animal <- subset %>%
  pivot_longer(cols = 3:9,
               names_to = "product",
               values_to = "CO2_person_year")

non_animal <- subset %>%
  pivot_longer(cols = 11:14,
               names_to = "product",
               values_to = "CO2_person_year")


```
### Final plot 1

```{r plot final animal}
  #plot1: animal products
#vector for highlighted countries
focus <- c("average","USA","Canada","Japan")

a2 <- animal %>% 
  ggplot(aes(product, CO2_person_year, group=country),show.legend = FALSE) +
  geom_line(color="grey",show.legend = FALSE) + 
  geom_line(aes(color = country),
            data = filter(animal, country %in% focus)) + 
  scale_color_viridis_d() +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "CO2/person/year for animal products",
       subtitle = "",
       x = "animal product",
       y = "Co2/person/year (in Kg)") +
  theme_minimal()
a2
ggplotly(a2, tooltip = c("country","product","CO2_person_year"))
```

Column {data-width=350}
-----------------------------------------------------------------------

### Initial plot 1

```{r plot v1 animal}
  #plot1: animal products
#draft

a1 <- animal %>% 
  ggplot(aes(product, CO2_person_year, group=country)) +
  geom_line(aes(color = country))
a1

```

### Version 2

```{r plot v2 animal}
  #plot1: animal products
a2 <- animal %>% 
  ggplot(aes(product, CO2_person_year, group=country)) +
  geom_line(aes(color = country), size = 1) +
  gghighlight(country == "average" |country == "USA" | country =="Canada"| 
                country =="Japan") +
  scale_color_viridis_d() +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "CO2/person/year for animal products",
       subtitle = "",
       x = "animal product",
       y = "Co2/person/year (in Kg)") +
  theme_minimal()
a2

```

Non-animal products
=========================
Column {data-width=650}
-----------------------------------------------------------------------

### Final plot 2

```{r plot final non-animal}
  #plot2: non-animal products
#final
na2 <- non_animal %>% 
  ggplot(aes(product, CO2_person_year, group=country),show.legend = FALSE) +
  geom_line(color="grey",show.legend = FALSE) +
  geom_line(aes(color = country),
            data = filter(non_animal, country %in% focus)) + 
  scale_color_viridis_d() +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "CO2/person/year for non-animal products",
       subtitle = "",
       x = "non-animal product",
       y = "Co2/person/year (in Kg)") +
  theme_minimal()
na2
ggplotly(na2, tooltip = c("country","product","CO2_person_year"))
```

Column {data-width=350}
-----------------------------------------------------------------------

### Initial plot 2 (non-animal)

```{r plot v1 non-animal}
na1 <- non_animal %>% 
  ggplot(aes(product, CO2_person_year, group=country)) +
  geom_line(aes(color = country))
na1
```

### Version 2

```{r plot non-animal v2}
  #plot2: non-animal products

na2 <- non_animal %>% 
  ggplot(aes(product, CO2_person_year, group=country)) +
    geom_line(aes(color = country), size = 1) +
  gghighlight(country == "average" |country == "USA" | country =="Canada"| 
                country =="Japan") +
  scale_color_viridis_d() +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "CO2/person/year for non-animal products",
       subtitle = "",
       x = "non-animal product",
       y = "Co2/person/year (in Kg)") +
  theme_minimal()
na2
```

Difference between animal and non-animal products {data-orientation=rows}
=========================
Row {vertical_layout: scroll}
-----------------------------------------------------------------------

### Difference between the CO2 production of animal product and non-animal product, over a year. A low value means that a larger proportion of the population feeds on plant products which have a better carbon emission footprint.


```{r plot final difference}

  #plot3: difference between animal and non-animal products
#final
d2 <- nadiff %>% 
  ggplot(aes(CO2_person_year, reorder(country, CO2_person_year))) +
  geom_col(aes(fill = CO2_person_year)) +
  scale_fill_continuous_diverging(
    "Green-Brown",
    rev = FALSE,
    mid = mean(initial$animal_nonanimal_difference, na.rm = TRUE)) +
  labs(title = "",
       subtitle = "",
       x = "Co2/person/year (Kg)",
       y = "") +
  theme_minimal()

d2
ggplotly(d2, tooltip = c("CO2_person_year"))
```

Column {data-width=650} vertical_layout: scroll
-----------------------------------------------------------------------

### Difference Mapped

Mapping the difference between the CO2 production of animal product and non-animal product, over a year. A low value means that a larger proportion of the population feeds on plant products which have a better carbon emission footprint.


The average difference between CO2 production of animal and non-animal products for the countries in this study is 700 kg. Countries with above average CO2 are shaded in brown;  those with below average CO2 are shaded in blue-green.
```{r draft map}

  #draft
#geographic plot work
#bring in map
country_data <- map_data("world") %>% 
  rename(country=region)

#left join initial and country_data by country
data_map<-full_join(initial, country_data)

#maybe use a map type other than albers; mercator?
#need outlines for all countries; full join? 
#now countries without data are filled grey

#plot of difference by country on map
m0 <- ggplot(data = data_map,
             mapping = aes(x = long, y = lat,
                           group = group, fill = animal_nonanimal_difference))
m1 <- m0 + geom_polygon(color = "gray90", size = 0.1) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) 
m2 <- m1 + scale_fill_continuous_diverging(
  "Green-Brown",
  rev = FALSE,
  mid = mean(initial$animal_nonanimal_difference, na.rm = TRUE)) +
  theme_minimal()

ggplotly(m2)
```